Multiple imputation of incomplete multilevel data using Heckman selection models

### 2022-11-11 1.Paper/Paper.Rmd
update references

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in section \ref{methodsn} we define the main steps of the proposed imputation method. In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

# The Heckman model
\label{heckman}
The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions. [@heckman1976]

@@

# Method
\label{methodsn}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.

@@

Initially, we use the copulat method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$. 
The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.

@@

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency. [@viechtbauer2005]
<!-- Insensitivity is not generally a good thing. I think you should either explain why it is a good thing here: "... it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "... a good balance between unbiasedness and efficiency. jajaj I think at certain point i used the translator to correct my writing and i did not notice that change unbiasedness by insensitive..But thanks God you have eagle eyes @@) .."  -->

@@

# Simulation study
\label{simulation}

@@

# An illustrative study
\label{example}
```{r, include=FALSE,cache = TRUE}
library(here)

@@

<!-- I see you wrote both parasitemia and parasitaemia. Choose one -->
To illustrate our proposed method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.  We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.

@@

<!-- The age at which children start going to school is not universal, so you need to mention this. done!! thanks -->
In terms of prevalence by age, there are no significant differences between methods (Figure \ref{figure:region_age}). The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions. Assuming that children start going to school after the age of 6, the results could be partly explained by the mobility of 2-6 year olds compared to school-age children, i.e. school-age children spend more time outdoors and travel more than younger children.
<!-- Can you explain in one sentence how mobility of children is related to prevalence here? OK-->

@@

# Discussion
\label{discussion}
We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
For clustered datasets, only imputation methods under the MAR mechanism had previously been proposed. Although imputation methods exist to handle MNAR they have only been designed for individual studies. This makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.   In this context, we proposed a new multiple imputation method to handle continuous and binary MNAR covariates specifically for a clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). thanks :D-->

@@

Further, the method can also be extended to other types of variables such as count or ordinal variables. Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by @ogundimu_collins19. 
<!-- Add reference to Ogundimu -->

>>>>

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{sec:heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in section \ref{sec:methods} we define the main steps of the proposed imputation method. In section \ref{sec:simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{sec:illustration} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion in section \ref{sec:discussion} about  the results, limitations and propose future extension of our method.

# The Heckman model
\label{sec:heckman}
The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions. [@heckman1976]

@@

# Method
\label{sec:methods}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.

@@

Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$. 
The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.

@@

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency. [@viechtbauer2005]
<!-- Insensitivity is not generally a good thing. I think you should either explain why it is a good thing here: "... it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "... a good balance between unbiasedness and efficiency. jajaj I think at certain point i used the translator to correct my writing and i did not notice that change unbiasedness by insensitive..But thanks God you have eagle eyes @@) .."  -->

@@

# Simulation study
\label{sec:simulation}

@@

# An illustrative study
\label{sec:illustration}
```{r, include=FALSE,cache = TRUE}
library(here)

@@

<!-- I see you wrote both parasitemia and parasitaemia. Choose one -->
To illustrate our proposed method, following the article by Rugnao et al. (2019),[@rugnao2019] we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.  We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.

@@

<!-- The age at which children start going to school is not universal, so you need to mention this. done!! thanks -->
In terms of prevalence by age, there are no significant differences between methods (Figure \ref{fig:region_age}). The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions. Assuming that children start going to school after the age of 6, the results could be partly explained by the mobility of 2-6 year olds compared to school-age children, i.e. school-age children spend more time outdoors and travel more than younger children.
<!-- Can you explain in one sentence how mobility of children is related to prevalence here? OK-->

@@

# Discussion
\label{sec:discussion}
We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
For clustered datasets, only imputation methods under the MAR mechanism had previously been proposed. Although imputation methods exist to handle MNAR they have only been designed for individual studies. This makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.   In this context, we proposed a new multiple imputation method to handle continuous and binary MNAR covariates specifically for a clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). thanks :D-->

@@

Further, the method can also be extended to other types of variables such as count or ordinal variables. Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu & Collins (2019).[@ogundimu_collins19]. 
<!-- Add reference to Ogundimu -->
